%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if it has changed)

\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

First and foremost, I would like to thank Dr. Jordi Nin for his advice and guidance throughout the entire development of this project, as well as his continued support to help me achieve my present and future goals. To him and all the professors I ever had, I am grateful of their will to teach and share their knowledge and sapience with all of us students; it is a great task you carry on your shoulders and for that you have my heartfelt gratitude.

Thanks to all inLab\textit{'ers} too. There, I have grown as an engineer and, most importantly, as a person, surrounded by amazing people. Thanks to Rosa, Albert and Jaume for their guidance, and thanks to all with whom I have had the most enthralling experiences at the faculty: Albert, Anna, Arnau, Carlos, Eli, Germán, Jaume, Joel, Maki, Marc, MJ, Natxo, Néstor, and sure I forget someone.

I am most thankful to my dearest friends Bernat, Borja, Laura, Natàlia, Pau; all these years have been amazing by your side. With you, life is brighter and the future is always awaiting on the horizon.

This project would not be without my family. All my gratitude and love is for my parents and my brother, who support me and raise me and teach me the most wonderful things. Thank you.

}
\clearpage % Start a new page
